The Role of Context in Spatial Region Identification

### 2013-09-13 scenarios/cdsr/trunk/papers/ACS2013/main.tex
jk fixed some typos

\title{CDSR Human Data/Modelling Paper}

@@

Both of these research questions address the fundamental claim that CDSRs are contextually defined. Question 1 addresses this claim by examining: (a) whether or not a CDSR is always applicable in a given enviroment (more bluntly, does a classroom always have a front or is this dependent on the configuration of objects in the room), and (b) the variation in CDSR placement and size within different context. Research question 2 addresses the contextual nature of CDSRs in terms of whether or not the variation in the location and extent of a CDSR systematically follow the variations  in the given context.

@@

We developed two experimental tasks that each address both research questions:
\begin{itemize}
	\item{Task 1:} subjects were presented with a visual representation of a spatial scene plus a linguistic description containing a CDSR and were tasked with drawing a polygon on the visual stimuli that defined the CDSR (or stated there was no such region).
	\item{Task 2:} subjects were presented with a visual representation of a spatial scene with a marked location and an accompanying spatial description and were asked to rate how well the location fitted with their understanding of the CDSR on a five point likert scale.
\end{itemize}

@@

The classroom stimuli consisted of pairs of floor plans and photos of classrooms with different chair configurations. For example, Figure \ref{fig:classroom-a} is a floor plan of a classroom where the chairs have been arranged in a circle and Figure \ref{fig:classroom-b} is the accompanying photo of this classroom. We developed four different classroom stimuli each with a different layout of desks: a circular layout, a U shaped layout (Figure \ref{fig:classroom-c} and Figure \ref{fig:classroom-d}), a typical row-based layout with the first row of chairs close to the expected front of the room (Figure \ref{fig:classroom-e}), and the same arrangement with the front row of chairs further away from the expected front of the room (Figure \ref{fig:classroom-f})\footnote{In the interest of saving space we do not show all the classroom photo stimuli}. To control for left-right bias effects, two versions of the U shaped layout, close rows layout and far rows layout were created by horizontally flipping the layouts. Consequently, there were seven classroom stimuli (circle, U $\times$ 2, close $\times$ 2, far $\times$ 2).  The two geographical map stimuli, Figures \ref{fig:bottleneck-a} and \ref{fig:bottleneck-b}, differed in the extent of the water.  Finally, the four military engagement stimuli, Figures \ref{fig:tanks-a}, \ref{fig:tanks-b}, \ref{fig:turrets-a}, \ref{fig:tanks-b} varied the functionality of the objects in the scene (tanks that could move versus fixed turret emplacements), the location of the objects in the scenes (were the tanks close to the bottom of the image or near the top), and the number and layout of the objects (the number and layout of the turrets were varied).

@@

\subsection{Task 1 - Results}
We use two methods to analyze the results of region drawing task.  First, we use interannotator agreement for two reasons: (1) if there is fair to good overall agreement, this is evidence that people change the regions they produce based systematically with context, (2) we expect less agreement in ambiguous contexts (e.g., the circle (a) and U-shaped (c) desk arrangements as well as the turret + 1 (l) military engagement configuration).  Second, we use the extent points of the regions to assess changes between trials.  That is, as the desks are moved back from stimuli (e) to (f), we expect the maximum x-coordinate of the region to increase.

%Kate: Question here: good agreement -> evidence that people change regions with context.  I think this needs to be reworded
% but want to talk to John about it during the telecon tomorrow first

@@

Table \ref{tab:stimuli-kappa-scores} lists the average pairwise $\kappa$ score for each of the visual stimuli used in the experiment and the number of no region responses for each stimuli.  It is clear from this table that the less common stimuli have the lowest $\kappa$ scores.  Each of the non-typical regions, the turret +1 arrangement (l), the u-shaped layouts (c and c-flpped), and the circle layout score have kappa scores of less than 0.35. Looking at the classroom results as a whole, these results support the hypothesis that the definition of CDSRs in a given context is effected by the location and orientation of objects in the environment.  This dependence is observed by the fact that just moving the desks into a circle configuration the CDSR for the front of the room no longer exists.

@@

Table \ref{tab:region-location-tests} compares the expected relationships between the geometric extent of the same region type between different contexts.  In five of the six comparisons, the expected relationships were found in human data supporting our claim that changes of context systematically change the extents of the regions.  The comparison of the safety area with tanks (j) versus turrets (k) is interesting because it provides some evidence that people are simulating actions of other agents in the scene.  That is, because the tanks can move and turrets cannot, the safe region extends higher on screen in the tanks context.  The case that did not demonstrate the expected the relationship was the safety region between the row of turrets (k) and turrets +1(l).  Our expectation was that people would reason that a lone turret would be no match for three tanks and indicate the safe region.  A possible explanation for this is that there is a domain expertise component to understanding contexts.

\subsection{Task 2 - Design} 
In task two, a location was marked on the visual stimuli (using a red dot) and the subject was asked to rank on a five point likert scale their judgement of how well the location matched the described spatial region, or select not applicable.  The location of these points was informed by the results of a pilot study, such that one of the domain's points, called the \textbf{sweetspot}, was located near the expected prototypical center of the described spatial region. In the classroom domain, the three points were located along the horizontal line that bisected the stimuli with P1 positioned on the left of the stimuli, P2 in the center and P3 on the right of the stimuli. This provides comparable data points between the trials where the classroom stimuli were horizontally flipped.  In the geography situations, P1 was positioned in the top left corner of the stimuli, P2 was the expected sweetspot and was located within the expected bottleneck region for both stimuli and P3 was located in the bottom right corner of the stimuli (see Figure \ref{fig:bottleneck-points}).  In the military engagement domain the three test points were located on the vertical line that bisected the stimuli: P1 was located near the enemy tanks, P3 was located with all friendly units between the it and the enemy tanks and P2 was located near the center of the image as follows. In stimuli (j,k) the point lay between all forces, in stimuli (i), it lay behind the friendly tanks, and in stimuli (l) , it behind the lone turret but in front of the row turrets.  P3 was the expected sweetspot in the military engagement domain but the rating of P2 should differ with context receiving the highest score in (i) and the lowest score in (k).

@@

Table \ref{tab:point-likert-scores} lists the average likert scores for the different points in each context. In all context, the expected sweetspot received the maximum average likert rating that was statistically significant using one-tailed two sample t-tests $(p<.05)$.  To control for left-right bias, we compared the ratings of sweetspots in the classroom stimuli (c,e and f) with their flipped variants.  There was no statistical difference between the stimuli and their flipped variants.  These likert ratings on individual stimuli supports the claim that people are able to agree on the locations of CDSRs with changing context.

@@

Another way of testing the relationship between the variation in context and the variation in definition of the CDSRs is to look at the differences in the ranking of the P2 point in the miliatary engagement domain between the pair of stimuli (i) and (j) and the pair of stimuli (k) and (l). In stimuli (i) the friendly tanks are positioned between the P2 point and the enemy tanks whereas in stimuli (j) the P2 point is in no-mans land. Given this change in the context one would expect a higher ranking for the P2 point in context (i) than in context (j). As the results in Table \ref{tab:point-likert-scores} show this is the case with average rating of P2  stimuli (i) being higher than that for (j). This higher rating is statiscially significant at a $95\%$ confidence level. Stimuli (k) and (l) present and analagous context with P2 being protected by an extra turret in stimuli (l) but in no-mans land in stimuli (k). Again, the average rating of P2 for stimuli (l) is higher than that of (k), although this difference is not statistically significant. Notice, that the rating of the sweetspot for the stimuli (i) and (l) remain quite high. This would indicate that the increase in the rating of the P2 points for these stimuli results from an increase in the area of the CDSR rather than a moving of the locus of the CDSR. Overall, the analysis of the ratings of the different sweetspots across the non-flipped and flipped stimuli and the variation in the ratings of the P2 points in the military domain stimuli point to the fact that both the location and the extent of a CDSR is contextually dependent.

@@

Spatial reasoning has been an active topic in AI and robotics for decades, CDSRs present a new problem which requires solutions beyond the current state-of-the-art. Regions are central to QSR work~\cite{Cohn:2001} and also in the area of cognitive robitics (see \emph{inter alia.} ~\cite{brenneretal07ijcai,kelleher/costello:09}). A fundamental decision when creating a computational model of a CDSR is the method used to represent its boundary. Given their context-dependent nature, boundary definitions based purely on single geometric shapes or a fixed number of points, seem unlikely to model the regions with sufficient accuracy. However, arbitrary boundaries can be defined using a sequence of geometric primitives (usually straight lines). This is an approach we followed in previous work \cite{Hawes:2012}, where we defined CDSRs using \textit{anchor points} as the vertices of the region polygon thereby linking the conceptual regions to perceived objects.  While satisfying the requirements of a cognitive systems solution, applying this model to our results of our experiment illustrates two limitations.  First, Regions are not always physically bounded by objects.  In the classroom examples, people frequently did not extend the front of the room all way to the desks.  Second, people consider dynamics part of the context, in the safety example, people did not indicate that it was safe behind the lone turret in front of the others. Finally, the connection between spatial regions and function has been explored in other robotics work, e.g. ~\cite{Karg:2012,Fasola:2013}, but these regions are based on fields anchored at single points in the environment or on a particular object. Therefore, they cannot account for the changes in context shown in this experiment which depend on multiple objects, agents, and simulation.  McLure and Forbus \cite{McLureForbus:2012} present a learning system that selects which encoding schemes work best for identifying geographic features (e.g., bottlenecks).  While this approach incorporates functional aspects of the situation, it is unclear how it would apply to the other situations in our experiment.

>>>>

\title{The Role of Context in Spatial Region Identification}

@@

Both of these research questions address the fundamental claim that CDSRs are contextually defined. Question 1 addresses this claim by examining: (a) whether or not a CDSR is always applicable in a given enviroment (e.g. does a classroom always have a front?), and (b) the variation in CDSR placement and size within different contexts. Research question 2 addresses the contextual nature of CDSRs in terms of whether or not the variation in the location and extent of a CDSR systematically follow the variations  in the given context.

@@

We developed two experimental tasks that each address both research questions:
\begin{itemize}
	\item{Task 1:} subjects were presented with a visual representation of a spatial scene plus a linguistic description containing a CDSR and were tasked with drawing a polygon on the visual stimuli that defined the CDSR (or stated there was no such region).
	\item{Task 2:} subjects were presented with a visual representation of a spatial scene with a marked location and an accompanying spatial description and were asked to rate how well the location fit with their understanding of the CDSR on a five point likert scale.
\end{itemize}

@@

The classroom stimuli consisted of pairs of floor plans and photos of classrooms with different chair configurations. For example, Figure \ref{fig:classroom-a} is a floor plan of a classroom where the chairs have been arranged in a circle and Figure \ref{fig:classroom-b} is the accompanying photo of this classroom. We developed four different classroom stimuli each with a different layout of desks: a circular layout, a U shaped layout (Figure \ref{fig:classroom-c} and Figure \ref{fig:classroom-d}), a typical row-based layout with the first row of chairs close to the expected front of the room (Figure \ref{fig:classroom-e}), and the same arrangement with the front row of chairs further away from the expected front of the room (Figure \ref{fig:classroom-f})\footnote{In the interest of saving space we do not show all the classroom photo stimuli}. To control for left-right bias effects, two versions of the U shaped layout, close rows layout and far rows layout were created by horizontally flipping the images. Consequently, there were seven classroom stimuli (circle, U $\times$ 2, close $\times$ 2, far $\times$ 2).  The two geographical map stimuli, Figures \ref{fig:bottleneck-a} and \ref{fig:bottleneck-b}, differed in the extent of the water.  Finally, the four military engagement stimuli, Figures \ref{fig:tanks-a}, \ref{fig:tanks-b}, \ref{fig:turrets-a}, \ref{fig:tanks-b} varied the functionality of the objects in the scene (tanks that could move versus fixed turret emplacements), the location of the objects in the scenes (were the tanks close to the bottom of the image or near the top), and the number and layout of the objects (the number and layout of the turrets were varied).

@@

\subsection{Task 1 - Results}
We use two methods to analyze the results of the region drawing task.  First, we use interannotator agreement for two reasons: (1) fair to good overall agreement, would provide evidence that people have a shared interpretation of CDSRs in the given contexts, (2) we expect less agreement in ambiguous contexts (e.g., the circle (a) and U-shaped (c) desk arrangements as well as the turret + 1 (l) military engagement configuration).  Second, we use the extent points of the regions to assess changes between trials.  That is, as the desks are moved back from stimuli (e) to (f), we expect the maximum x-coordinate of the region to increase showing that region extent varies systematically with context changes.

%Kate:  (1) if there is fair to good overall agreement, this is evidence that people change the regions they produce based systematically with %context,  I changed this SUBSTANTIALLY to something that I think sounds better, but makes a different argument.  Let me know if it is ok.  Really %agreement would indicated shared understanding while variation between contexts would lend support to the importance of context.  Not sure I %captured this

@@

Table \ref{tab:stimuli-kappa-scores} lists the average pairwise $\kappa$ score for each of the visual stimuli used in the experiment and the number of no region responses for each stimuli.  It is clear from this table that the less common stimuli have the lowest $\kappa$ scores.  Each of the non-typical regions, the turret +1 arrangement (l), the u-shaped layouts (c and c-flpped), and the circle layout have kappa scores of less than 0.35. Looking at the classroom results as a whole, these results support the hypothesis that the definition of CDSRs in a given context is effected by the location and orientation of objects in the environment.  This dependence is observed by the fact that just moving the desks into a circle configuration the CDSR for the front of the room no longer exists.

@@

Table \ref{tab:region-location-tests} compares the expected relationships between the geometric extent of a CDSR in different contexts.  In five of the six comparisons, the expected relationships were found, supporting our claim that changes of context systematically change the perceived extent of a CDSR.  The comparison of the safety area with tanks (j) versus turrets (k) is interesting as it provides some evidence that people are simulating actions of other agents in the scene.  Because the tanks can move and turrets cannot, the safe region extends higher on screen in the tanks context.  The case that did not demonstrate the expected the relationship was the safety region between the row of turrets (k) and turrets +1(l).   A possible explanation for this is that there is a domain expertise component to understanding contexts.

%Kate: I took out "Our expectation was that people would reason that a lone turret would be no match for three tanks and indicate the safe region. " because it didn't make sense to me.

\subsection{Task 2 - Design} 
In task two, a location was marked on the stimuli (using a red dot) and the subject was asked to rank on a five point likert scale how well the location matched the given CDSR, or select not applicable.  The location of these points was informed by the results of a pilot study, such that one of the domain's points, called the \textbf{sweetspot}, was located near the expected prototypical center of the CDSR. In the classroom domain, the three points were located along the horizontal line that bisected the stimuli with P1 positioned on the left of the stimuli, P2 in the center and P3 on the right of the stimuli. This provides comparable data points between the trials where the classroom stimuli were horizontally flipped.  In the geography situations, P1 was positioned in the top left corner of the stimuli, P2 was the expected sweetspot and was located within the expected bottleneck region for both stimuli and P3 was located in the bottom right corner of the stimuli (see Figure \ref{fig:bottleneck-points}).  In the military engagement domain the three test points were located on the vertical line that bisected the stimuli: P1 was located near the enemy tanks, P3 was located with all friendly units between the it and the enemy tanks and P2 was located near the center of the image as follows. In stimuli (j,k) the point lay between all forces, in stimuli (i), it lay behind the friendly tanks, and in stimuli (l) , it behind the lone turret but in front of the row turrets.  P3 was the expected sweetspot in the military engagement domain but the rating of P2 should differ with context receiving the highest score in (i) and the lowest score in (k).

@@

Table \ref{tab:point-likert-scores} lists the average likert scores for the different points in each context. In all contexts, the expected sweetspot received the maximum average likert rating that was statistically significant using one-tailed two sample t-tests $(p<.05)$.  To control for left-right bias, we compared the ratings of sweetspots in the classroom stimuli (c,e and f) with their flipped variants.  There was no statistical difference between the stimuli and their flipped variants.  These likert ratings on individual stimuli support the claim that people are able to agree on the locations of CDSRs with changing context.

@@

Another way of testing the relationship between the variation in context and the variation in definition of the CDSRs is to look at the differences in the ranking of the P2 point in the military domain between the stimuli (i) and (j) and the stimuli (k) and (l). In stimuli (i) the friendly tanks are positioned between the P2 point and the enemy tanks whereas in stimuli (j) the P2 point is in no-mans land. Given this change in context one would expect a higher ranking for the P2 point in context (i) than in context (j). As the results in Table \ref{tab:point-likert-scores} show this is the case with this higher rating statistically significant $(p<.05)$. Stimuli (k) and (l) present an analagous context with P2 being protected by an extra turret in stimuli (l) but in no-mans land in stimuli (k). Again, the average rating of P2 for stimuli (l) is higher than that of (k), although this difference is not statistically significant. Notice, that the rating of the sweetspot for the stimuli (i) and (l) remain quite high. This would indicate that the increase in the rating of the P2 points for these stimuli results from an increase in the area of the CDSR rather than a moving of the locus of the CDSR. Overall, the analysis of the ratings of the different sweetspots across the non-flipped and flipped stimuli and the variation in the ratings of the P2 points in the military domain stimuli point to the fact that both the location and the extent of a CDSR is contextually dependent.

@@

Spatial reasoning has been an active topic in AI and robotics for decades, CDSRs present a new problem which requires solutions beyond the current state-of-the-art. Regions are central to QSR work~\cite{Cohn:2001} and also in the area of cognitive robitics (see \emph{inter alia.} ~\cite{brenneretal07ijcai,kelleher/costello:09}). A fundamental decision when creating a computational model of a CDSR is the method used to represent its boundary. Given their context-dependent nature, boundary definitions based purely on single geometric shapes or a fixed number of points, seem unlikely to model the regions with sufficient accuracy. However, arbitrary boundaries can be defined using a sequence of geometric primitives (usually straight lines). This is an approach we followed in previous work \cite{Hawes:2012}, where we defined CDSRs using \textit{anchor points} as the vertices of the region polygon thereby linking the conceptual regions to perceived objects.  While satisfying the requirements of a cognitive systems solution, applying this model to the results of our experiment illustrates two limitations.  First, Regions are not always physically bounded by objects.  In the classroom examples, people frequently did not extend the front of the room all way to the desks.  Second, people consider dynamics of a context, in the safety example, people did not indicate that it was safe behind the lone turret in front of the others. Finally, the connection between spatial regions and function has been explored in other robotics work, e.g. ~\cite{Karg:2012,Fasola:2013}, but these regions are based on fields anchored at single points in the environment or on a particular object. Therefore, they cannot account for the changes in context shown in this experiment which depend on multiple objects, agents, and simulation.  McLure and Forbus \cite{McLureForbus:2012} present a learning system that selects which encoding schemes work best for identifying geographic features (e.g., bottlenecks).  While this approach incorporates functional aspects of the situation, it is unclear how it would apply to the other situations in our experiment.
